Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization

### 2023-10-13 manuscript.Rmd
add more references on deviations after preregistration
cite computational reproducibility

This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration, to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
Thus, insisting on equating preregistration and confirmation has led to the criticism that, all things considered, preregistration is actually harmful and neither sufficient nor necessary for doing good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].

@@

Communicating clearly how authors of a scientific report collected their data and consequently analyzed it to arrive at the evidence they present is crucial for judging the theoretical risk they took.
Preregistrations are ideal for communicating just that because any description after the fact is prone to be incomplete.
For instance, the authors could have opted for selective reporting, that is, they decided to exclude a number of analytic strategies they tried out.
That is not to say that every study that was not-preregistered was subjected to practices of questionable research practices.
The point is that we cannot exclude it with certainty.
This uncertainty is drastically reduced if the researchers have described what they intended to do beforehand and then report that they did exactly that.
In that case, readers can be certain they received a complete account of the situation.
They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extent than if the study would not have been preregistered.
The remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analyses, etc.
In any case, a well-written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
Therefore, a study that perfectly adhered to its preregistration will resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
Crucially, perfect means here that the theoretical risk can be judged with low uncertainty, not that the theoretical risk is necessarily high.

@@

Theoretical risk may conceptually be related to the very fruitful conceptual framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
However, there are crucial differences.
First, we are not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
Second, severity considerations are mainly post-data.
Only after the evidence was observed can the severity with which we can rule out alternatives be calculated.
Third, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .9999.
However, they also share properties, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].

>>>>

This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration, to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
The latter is a frequent occurrence and, perhaps more worryingly, often remains undisclosed [@akkerEffectivenessPreregistrationPsychology2023; @claesenComparingDreamReality2021].
Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
Thus, insisting on equating preregistration and confirmation has led to the criticism that, all things considered, preregistration is actually harmful and neither sufficient nor necessary for doing good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].

@@

Communicating clearly how authors of a scientific report collected their data and consequently analyzed it to arrive at the evidence they present is crucial for judging the theoretical risk they took.
Preregistrations are ideal for communicating just that because any description after the fact is prone to be incomplete.
For instance, the authors could have opted for selective reporting, that is, they decided to exclude a number of analytic strategies they tried out.
That is not to say that every study that was not-preregistered was subjected to practices of questionable research practices.
The point is that we cannot exclude it with certainty.
This uncertainty is drastically reduced if the researchers have described what they intended to do beforehand and then report that they did exactly that.
In that case, readers can be certain they received a complete account of the situation.
They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extent than if the study would not have been preregistered.

The remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analyses, etc.
To further reduce the uncertainty about theoretical risk, researchers must therefore publish code and ideally data.
After all, computational reproducibility is only possible if the data analytic procedure was communicated clearly enough to allow others to retrace the computational steps [@peikertReproducibleDataAnalysis2021].

In any case, a well-written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
Therefore, a study that perfectly adhered to its preregistration will resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
Crucially, perfect means here that the theoretical risk can be judged with low uncertainty, not that the theoretical risk is necessarily high.

@@

Theoretical risk may conceptually be related to the very fruitful conceptual framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
However, there are crucial differences.
First, we are not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
This is important because, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .9999.
Second, severity considerations are mainly post-data.
Only after the evidence was observed can the severity with which we can rule out alternatives be calculated.
However, they also share properties, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].
